DATA 621 - Business Analytics and Data Mining

### 2020-11-24 HW5/HW5-Draft.Rmd
TOC
NB Model 2 & ML Model 1

```{r loadData, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(knitr)
library(MASS)
library(caret)
library(corrplot)
library(data.table)

@@

@@

## Multiple Linear Regrssion Model # 1

@@

### Multiple Linear Regression Model 1

@@

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
<!-- Assignment -->
```{r loadData}
```
<!-- Data Exploration -->
```{r missingVal}
```
```{r statsN}
```
```{r graphsD}
```
```{r graphsH}
```
```{r corrgram}
```
```{r targetH}
```
<!-- Data Preparation -->
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r model1impute}
```
```{r m1postImputeD}
```
```{r m1postImputeH}
```
<!-- Build Models -->
```{r m1PTrain}
```
```{r m1PTable}
```
```{r m1PFitCheck}
```
```{r m1PvarImp}
```
```{r m1NBTrain}
```
```{r m1NBTable}
```
```{r m1NBvarImp}
```
```{r m1PNB}
```
<!-- Other prep and build goes here -->
<!-- Select Models -->
```{r compTable}
```
```{r m1TestProcess}
```
```{r m1PTest}
```
<!-- Poisson 2 Test goes here -->
```{r m1NBTest}
```
<!-- Remaining Test sets go here -->
```{r modelTestResults}
```

>>>>

```{r loadData, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(knitr)
library(MASS)
library(caret)
library(corrplot)
library(data.table)
library(dplyr)
library(mice)

@@

## Multiple Linear Regression 1 & Negative Binomial Model 2
### Missing Data
These models will deal with missing numerical data by imputing the NA's
using the Multivariate imputation by chained equations (MICE) method. Multiple
imputation involves creating multiple predictions for each missing value, helping 
to account for the uncertainty in the individual imputations.

```{r model2impute}
set.seed(123)

# ML1 & NB2 data transformation
# MICE imputation on train and test set
m2train <- complete(mice(data = trnSet %>% select(-INDEX),
                         method = "pmm", print = FALSE), 3)

m2test <- complete(mice(data = tstSet %>% select(-INDEX),
                        method = "pmm", print = FALSE), 3)

```

@@

#### Baseline Model 
We will start with a simple negative binomial regression model to serve as a 
baseline. This includes all variables in the dataset.

```{r m2NBTrainBase}

m2NBBase <- glm.nb(TARGET ~ ., data = m2train)
knitr::kable(summary(m2NBBase)$coefficients, digits = 3L,
             caption = 'Negative Binomial Regression Model 2 - Base Negative Binomial Output')

```
Immediately, we can see a few variables that do not meet the 0.05
p-value threshold.

#### Enhanced Model
Backwards stepwise regression is performed and the result is a model with the 
following variables *removed*: `FixedAcidity`, `CitricAcid`, `ResidualSugar`, 
and `Density`. Because the `Sulphates` variable is very close to the 0.05
p-value threshold of significance, we will remove it as well.

```{r m2NBTrain}

m2NBFit <- glm.nb(TARGET ~ VolatileAcidity + Chlorides +
                FreeSulfurDioxide + TotalSulfurDioxide + pH + Alcohol + 
                LabelAppeal + AcidIndex + STARS, data = m2train)

knitr::kable(summary(m2NBFit)$coefficients, digits = 3L,
             caption = 'Negative Binomial Regression Model 2 - Final Negative Binomial Output')

```

#### Coefficient Discussion
The intercept of the final model tells us that there is typically ~1 case 
of wine purchased. Some things to note:

* Stars and label appeal contribute the most positively to the number of
cases of wine purchased. Intuitively, this makes sense - the higher the
rating and the prettier the bottle, the more cases we would expect to be
sold. 
* Free Sulfur Dioxide and Total Sulfur Dioxide have a very tiny effect on
the number of cases sold (rounded to 3 decimal places, the impact is 0!)
* All of the remaining variables except alcohol have a negative effect on 
the number of cases purchased.

## Multiple Linear Regrssion Model # 1
One drawback to traditional multiple linear regression models is that the 
forecasted counts are not integers. We will deal with this by rounding the
predictions to the nearest whole number.

#### Baseline Model
We will start with a simple linear model to serve as a baseline. This includes
all variables in the dataset.

```{r m1MLTrainBase}

m1MLBase <- lm(TARGET ~ ., data = m2train)
knitr::kable(summary(m1MLBase)$coefficients, digits = 3L,
             caption = 'Multiple Regression Model 1 - Base Linear Regression Output')

```
Most of the variables in the dataset are significant, but we can immediately
see a few that are above the 0.05 p-value threshold.

#### Enhanced Model
Backwards stepwise regression is performed and the result is a model with the 
following variables *removed*: `FixedAcidity`, `CitricAcid`, `ResidualSugar`, 
and `Density`. 
```{r m1MLTrain}

m1MLFit <- lm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide + TotalSulfurDioxide + pH +
             Sulphates + Alcohol + LabelAppeal + AcidIndex + STARS, data = m2train)
knitr::kable(summary(m1MLFit)$coefficients, digits = 3L,
             caption = 'Multiple Regression Model 1 - Final Linear Regression Output')

```

#### Coefficient Discussion
The intercept of the final model tells us that there are typically ~3 cases 
of wine purchased. Some things to note:

* As expected, the number of stars has the largest, positive impact on the 
number of cases sold. This makes sense -- the higher the rating of a wine,
the more of it we would expect to be purchased. 
* Label appeal has the second largest impact on the number of cases sold. 
They say one shouldn't judge a book by its cover (or in this case, a wine
by its label), but it does appear to have quite a significant impact on
sales. 
* Alcohol also has a small positive impact on the number of cases purchased. 
This means that the more alcohol a wine has, the more cases of it will be
purchased. 
* Acid Index, Sulphates, pH, Chlorides, and VolatileAcidity all have a
negative impact on the number of cases sold.

@@

```{r m2NBTest}
# NB2: Test Model
m2NBtstPred <- round(predict(m2NBFit, newdata = m2test))
compTable[4, 2] <- m2NBRMSE <- RMSE(m2NBtstPred, m2test$TARGET)
compTable[4, 3] <- m2NBMAE <- MAE(m2NBtstPred, m2test$TARGET)
compTable[4, 4] <- m2NBFitAIC <- AIC(m2NBFit)
```

**Negative Binomial Model 2** has an RMSE of
`r prettyNum(m2NBRMSE, digits = 2L)`, an MAE of
`r prettyNum(m2NBMAE, digits = 2L)`, and a training AIC of
`r prettyNum(m2NBFitAIC, digits = 2L)`.

### Multiple Linear Regression Model 1
```{r m1MLTest}
# ML1: Test Model
m1MLtstPred <- round(predict(m1MLFit, newdata = m2test))
compTable[5, 2] <- m1MLMSE <- RMSE(m1MLtstPred, m2test$TARGET)
compTable[5, 3] <- m1MLMAE <- MAE(m1MLtstPred, m2test$TARGET)
compTable[5, 4] <- m1MLFitAIC <- AIC(m1MLFit)

```

**Multiple Linear Regression Model 1** has an RMSE of
`r prettyNum(m1MLMSE, digits = 2L)`, an MAE of
`r prettyNum(m1MLMAE, digits = 2L)`, and a training AIC of
`r prettyNum(m1MLFitAIC, digits = 2L)`.

@@

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
<!-- Assignment -->
```{r loadData}
```
<!-- Data Exploration -->
```{r missingVal}
```
```{r statsN}
```
```{r graphsD}
```
```{r graphsH}
```
```{r corrgram}
```
```{r targetH}
```
<!-- Data Preparation -->
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r model1impute}
```
```{r m1postImputeD}
```
```{r m1postImputeH}
```
```{r model2impute}
```
<!-- Build Models -->
```{r m1PTrain}
```
```{r m1PTable}
```
```{r m1PFitCheck}
```
```{r m1PvarImp}
```
```{r m1NBTrain}
```
```{r m1NBTable}
```
```{r m1NBvarImp}
```
```{r m1PNB}
```
```{r m2NBTrainBase}
```
```{r m2NBTrain}
```
```{r m1MLTrainBase}
```
```{r m1MLTrain}
```
<!-- Other prep and build goes here -->
<!-- Select Models -->
```{r compTable}
```
```{r m1TestProcess}
```
```{r m1PTest}
```
<!-- Poisson 2 Test goes here -->
```{r m1NBTest}
```
```{r m2NBTest}
```
```{r m1MLTest}
```
<!-- ML 2 Test goes here -->
```{r modelTestResults}
```
